---
title: '`Course Project` Regression Models'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
params:
  author: 'Anderson H Uyekita'
  course: 'Regression Models'
  course_url: 'https://www.coursera.org/learn/regression-models'
  specialization: 'Data Science: Foundations using R Specialization'
  specialization_url: 'https://www.coursera.org/specializations/data-science-foundations-r'
  instructor: 'Brian Caffo'
  course_start: !r lubridate::ymd('2022/07/05', tz = 'America/Sao_Paulo')
  course_finish: !r lubridate::today(tzone = 'America/Sao_Paulo')
  week: "Week 4"
  slug: 'Week%204'
  rpubs: 'https://rpubs.com/AndersonUyekita/regression-models_reproducible-research'
  project_repo: 'https://github.com/AndersonUyekita/regression-models_course-project'
  codebook: 'https://github.com/AndersonUyekita/regression-models_course-project/blob/master/codebook.md'
  instructions: 'https://github.com/AndersonUyekita/regression-models_course-project/blob/master/instructions.md'
always_allow_html: true
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo=TRUE,cache=TRUE,message=FALSE,warning=FALSE)
library(rmarkdown)
library(shiny)
library(datasets)
library(explore)
library(dplyr)
```

`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1F30E; Rpubs: [Interactive Document](%s)", params$rpubs)}else{sprintf(fmt = "* &#x1F4E6; Github Repository: [Static Document](%s){target='_blank' rel='noopener'}", params$project_repo)}`
`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1F4CB; Instructions: [Project Instructions](%s)", "./instructions.md")}else{sprintf(fmt = "* &#x1F4CB; Instructions: [Project Instructions](%s){target='_blank' rel='noopener'}", params$instructions)}`
`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1F4C4; README: [README.md](%s)", "./README.md")}else{sprintf(fmt = "* &#x1F4C4; README: [README.md](%s){target='_blank' rel='noopener'}", params$project_repo)}`

--------------------------------------------------------------------------------

## Codebook

According to the [R Documentation][1]

[1]: https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html

### Description

> Henderson and Velleman (1981) comment in a footnote to Table 1: ‘Hocking [original transcriber]'s noncrucial coding of the Mazda's rotary engine as a straight six-cylinder engine and the Porsche's flat engine as a V engine, as well as the inclusion of the diesel Mercedes 240D, have been retained to enable direct comparisons to be made with previous analyses.’

### Dimensions

* Observations: `r dim(datasets::mtcars)[1]`
* Variables: `r dim(datasets::mtcars)[2]`

### `str`

```{r,echo=FALSE}
str(datasets::mtcars)
```

### Variables Description

* `mpg` : Miles/(US) gallon
* `cyl` : Number of cylinders
* `disp` : Displacement (cu.in.)
* `hp` : Gross horsepower
* `drat` : Rear axle ratio
* `wt` : Weight (1000 lbs)
* `qsec` : 1/4 mile time
* `vs` : Engine (0 = V-shaped, 1 = straight)
* `am` : Transmission (0 = automatic, 1 = manual)
* `gear` : Number of forward gears
* `carb` : Number of carburetors

### Summary

```{r,echo=FALSE}
summary(datasets::mtcars)
```

### Source

> Henderson and Velleman (1981), Building multiple regression models interactively. Biometrics, 37, 391–411.

### Exploratory

```{r,echo=FALSE}
mtcars %>% 
  select(mpg, cyl, disp, hp, gear) %>% 
  explore_all(target = gear)
```

```{r,echo=FALSE}
mtcars %>% 
  select(drat, wt, qsec, vs, gear) %>% 
  explore_all(target = gear)
```

```{r,echo=FALSE}
mtcars %>% 
  select(am, carb,vs, gear) %>% 
  explore_all(target = gear)
```