---
title: '`Course Project` Regression Models'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
  github_document: default
params:
  author: 'Anderson H Uyekita'
  course: 'Regression Models'
  course_url: 'https://www.coursera.org/learn/regression-models'
  specialization: 'Data Science: Statistics and Machine Learning Specialization'
  specialization_url: 'https://www.coursera.org/specializations/data-science-statistics-machine-learning'
  instructor: 'Brian Caffo'
  course_start: !r lubridate::ymd('2022/07/05', tz = 'America/Sao_Paulo')
  course_finish: !r lubridate::ymd('2022/07/09', tz = 'America/Sao_Paulo')
  week: "Week 4"
  slug: 'Week%204'
  rpubs: 'https://rpubs.com/AndersonUyekita/course-project_regression-models'
  project_repo: 'https://github.com/AndersonUyekita/regression-models_course-project'
  codebook: 'https://github.com/AndersonUyekita/regression-models_course-project/blob/master/codebook.md'
  instructions: 'https://github.com/AndersonUyekita/regression-models_course-project/blob/master/instructions.md'
always_allow_html: true
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo=TRUE,cache=TRUE,message=FALSE,warning=FALSE)
library(rmarkdown)
```

`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1F30E; Rpubs: [Interactive Document](%s)", params$rpubs)}else{sprintf(fmt = "* &#x1F4E6; Github Repository: [Static Document](%s){target='_blank' rel='noopener'}", params$project_repo)}`


--------------------------------------------------------------------------------

## Brief Analysis about Manual and Automatic Transmissions

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/AndersonUyekita/regression-models_course-project/main?urlpath=rstudio)

## Executive Summary

The data analysis process in this document has identified that manual vehicles have better performance than automatics concerning miles per gallon (`mpg`). Furthermore, based on this study's linear regression model, the difference has reached 7.24 miles per gallon based on the average weights of automatic and manual cars (See Figure 4 in APPENDIX Section A4.), which is a significant number.

```{r loading_packages,echo=FALSE,warning=FALSE,message=FALSE}
# Loading libraries
library(ggplot2)
library(tidyverse)
library(magrittr)
library(datasets)
library(explore)
library(kableExtra)
library(gridExtra)

# Force results to be in English
base::invisible(Sys.setlocale("LC_ALL", "English.utf8"))

# Set seed to turn it reproducible
base::set.seed(2022)

knitr::opts_chunk$set(fig.pos = "!H", out.extra = "")
```

## 1. Introduction

The _Motor Trend Car Road Test_ has evaluated 32 automobiles varying from 1973-74 models, and this study comprises fuel consumption, horsepower, weight, and other aspects. Based on it, the present publication aims to answer two questions:

1. "Is an automatic or manual transmission better for MPG?"
2. "Quantify the MPG difference between automatic and manual transmissions."

## 2. Requeriments and Settings

If you are interested in reproducing this study, please visit the [Github repository][github_url] to have access to the raw document.

[github_url]: https://github.com/AndersonUyekita/regression-models_course-project

## 3. Loading Data and EDA

```{r loading-data,echo=FALSE}
# Loading the mtcars dataset.
df_cars <- datasets::mtcars

# Data manipulation
df_cars <- df_cars %>%
    mutate(
        vs = factor(vs, labels = c("V", "S")),
        am = factor(am, labels = c("automatic", "manual")),
        cyl  = ordered(cyl),
        gear = ordered(gear),
        carb = ordered(carb))
```

The adjusted data frame has 32 observations with no `NA` values, divided into 6 (six) numeric variables (no one is standardized or scaled) and 5 (five) categorical variables. For more exploratory details, please, find them in APPENDIX section A1. For more information about the variables descriptions, please, see it on the [R Documentation website][mtcars_documentation].

[mtcars_documentation]:https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html

```{r t-test,echo=FALSE}
# Testing the H0 hypothesis.
#
# H0 : mu_automatic  = mu_manual
# H1 : mu_automatic != mu_manual
#
# Due to the size of the sample, I have used the t-test.
#
t.test(subset(x = df_cars, subset = am == "automatic", select = mpg),
       subset(x = df_cars, subset = am == "manual", select = mpg),
       paired = FALSE,
       alternative = "two.sided") -> c3_t.test
```

Due to Figure 2 in section A2 from APPENDIX, I have tested the hypothesis that the average consumption from automatic and manual vehicles is equal. In other words, $H_O: \mu_{automatic} = \mu_{manual}$, the p-value is `r round(c3_t.test$p.value, 4)*100`% which is way less than alpha (5%). For this reason, the $H_O$ was **Rejected**, which means the averages of automatic and manual transmissions are from different populations.

## 4. Model Selection

The model selection approach used for this project is based on the Week 3 videos and the Chapter Multiple variables and model selection from Regression Models for Data Science in R book.

### 4.1. Base line model

$$lm(formula  = mpg \sim am, data = mtcars)$$

```{r baseline-model,echo=FALSE,message=FALSE,warning=FALSE}
# Baseline model using only transmission.
fit1 <- lm(data = df_cars, formula = mpg ~ am)

# Storing the summary info.
summary_fit1 <- summary(fit1)
```

The baseline model is the Ordinary Linear Regression, and this model uses only transmission (`am`) to explain the consumption in miles per gallon (`mpg`). From the hypothesis tested in Section 3, this baseline model has identified that manual transmission performs better than automatic ones. On average, manual transmission yields `r round(summary_fit1$coefficients[1,1], 2) + round(summary_fit1$coefficients[2,1], 2)` miles per gallon. On the other hand, automatic transmission yield `r round(summary_fit1$coefficients[1,1], 2)` miles per gallon. The difference is `r round(summary_fit1$coefficients[2,1], 2)` miles per gallon.

### 4.2. Analysis of Variance

Using the `anova()` function, it was possible to run several combinations, reaching the final model using am, wt, and interaction between am and wt.

$$lm(formula  = mpg \sim am + wt + am \cdot wt, data = mtcars)$$

```{r,echo=FALSE,warning=FALSE,message=FALSE}
# Baseline
fit1 <- lm(data = df_cars, formula = mpg ~ am)

# Adding wt
fit2 <- update(fit1, formula = mpg ~ am + wt)

# Adding relationship between am and wt.
fit3 <- update(fit1, formula = mpg ~ am + wt + am * wt)

# Creating the final model.
fit_final <- fit3;
```

I have decided to use a simpler model due to the parsimony. The R2 adjusted has reached `r round(summary(fit_final)$adj.r.squared,4)*100`%. All p-values of the model are below alpha (5%).

The linear model coefficients: 

```{r,fig.pos="!H",echo=FALSE}


summary(fit_final)$coeff %>% kbl(booktabs = TRUE) %>% kable_styling(latex_options = c("striped","hold_position"))
```

## 5. Residual Analysis

```{r,echo=FALSE,warning=FALSE,message=FALSE}
c5.test <- shapiro.test(resid(fit_final))
```

Due to the low number of observations, below 50, I have used the Shapiro-Wilk test to ensure the residual's normality. The p-value obtained from this test was `r round(c5.test$p.value,4)*100`%, sufficient to reject the null hypothesis and proving the residual's normality.

The residual analysis will be based on Figure 3 in APPENDIX Section A3.. This figure aims to corroborate the following explanations:

* **Residual vs. Fitted:** The residual bounces around zero, which suggests an excellent linear relationship. Also, the residual is a homogeneous spread on the plot showing that the variance is "constant" in addition, I have identified no outlier.
* **Normal Q-Q:** Most of the data is around the line, except in the top right of the chart. I did not identify any issue related to non-normality.
* **Scale-Location:** There are no patterns in the points, the data stay between a fixed band, and the red line is almost constant. From those characteristics, the residual is considered homoscedasticity.
* **Residual vs. Leverage:** No points above the dotted line means there is no influential high point.

## 6. Conclusion

Considering the baseline model, the `am` variable can explain `r round(summary(fit1)$adj.r.squared,4)*100`% of the miles per gallon. However, In the final model, the percentage of variance explained by the model rises to `r round(summary(fit_final)$adj.r.squared,4)*100`% with the inclusion of two more predictors (`wt` and interaction of `wt` and `am`).

* **Final model:**

$$mpg=
\begin{cases}
\text{Automatic vehicle (am = 0)} \implies \beta_0 + \beta_2 \cdot wt = 31.42 -3.79 \cdot wt \\
\text{Manual vehicle (am = 1)} \implies (\beta_0 + \beta_1) + (\beta_2 + \beta_3) \cdot wt = 46.29-9.08 \cdot wt
\end{cases}$$

<!-- Adding a Page Break to starting a new APPENDIX page -->
\newpage

\center

## APPENDIX

\raggedright

### A1. Exploratory Data Analysis

```{r,fig.height=5,echo=FALSE}
df_cars %>% 
    select(-am) %>%
    explore_all(target = mpg, ncol = 4)
```

\center

**Figure 1** -- Exploratory Data Visualization

\raggedright

### A2. Miles per Galon (`mpg`) vs Transmission (`am`)

```{r eda,fig.height=2,echo=FALSE,warning=FALSE,message=FALSE}
# Creating a plot of MPG and AM.
df_cars %>% 
    ggplot(aes(x = am, y = mpg, fill = am)) + 
    geom_boxplot() +
    theme(legend.position="none")-> fig2.1

# Creating a facet grid of MPG and AM.
df_cars %>% 
    ggplot(aes(x = am, y = mpg, fill = am)) + 
    geom_violin() + 
    geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 0.5) +
    facet_grid(cols = vars(gear)) +
    theme(legend.position="none") -> fig2.2

# Combining the above two graphs into one.
grid.arrange(fig2.1, fig2.2, ncol=2, widths=c(1.5, 4))
```

\center

**Figure 2** -- Fuel Consumption divided into Transmission and number of Gears.

\raggedright

### A3. Residuals

```{r residuals,fig.height=6,echo=FALSE,warning=FALSE,message=FALSE}
par(mfrow = c(2, 2))
plot(fit_final)
```

\center

**Figure 3** -- Residuals.

\raggedright

### A4. Regression Lines

```{r,echo=FALSE,fig.height=6}
# Calculating the MPG for Automatic Cars based on average WT
car_0 <- 31.416055 - 3.785907 * mean(subset(x = df_cars, am == "automatic", select = wt)$wt)

# Calculating the MPG for Manual Cars based on average WT
car_1 <- 46.294477 - 9.084268 * mean(subset(x = df_cars, am == "manual", select = wt)$wt)

# Plotting.
df_cars %>%
    
    # A scatter plot.
    ggplot(aes(x = wt, y = mpg)) + 
    geom_point(aes(color = am)) + 
    
    # AUTOMATIC: Adding Automatic regression line.
    geom_abline(slope = - 3.785907,
                intercept = 31.416055,
                color = "#F8766D") + 
    
    # AUTOMATIC: Adding horizontal line until weight average
    geom_segment(data = data.frame("mpg" = car_0,
                                   "wt" = mean(subset(x = df_cars, am == "automatic", select = wt)$wt)),
                 aes(xend = -Inf, yend = car_0),
               color = "#F8766D",
               lty = 2) + 
    
    # AUTOMATIC: Adding vertical line until mpg average
    geom_segment(data = data.frame("mpg" = car_0,
                                   "wt" = mean(subset(x = df_cars, am == "automatic", select = wt)$wt)),
                 aes(xend = mean(subset(x = df_cars, am == "automatic", select = wt)$wt), yend = -Inf),
               color = "#F8766D",
               lty = 2) +     
    
    # MANUAL: Adding Automatic regression line.
    geom_abline(slope = - 9.084268,
                intercept = 46.294477,
                color = "#00BFC4") +
    
    # MANUAL: Adding horizontal line until weight average
    geom_segment(data = data.frame("mpg" = car_1,
                                   "wt" = mean(subset(x = df_cars, am == "manual", select = wt)$wt)),
                 aes(xend = -Inf, yend = car_1),
               color = "#00BFC4",
               lty = 2) + 

    # MANUAL: Adding vertical line until mpg average
    geom_segment(data = data.frame("mpg" = car_1,
                                   "wt" = mean(subset(x = df_cars, am == "manual", select = wt)$wt)),
                 aes(xend = mean(subset(x = df_cars, am == "manual", select = wt)$wt), yend = -Inf),
               color = "#00BFC4",
               lty = 2) +   
    
    # Adding labels of mog average for each kind of transmission.
    geom_label(data = data.frame("wt" = c(0,0), "mpg" = c(car_0, car_1)),
              aes(x = wt,
                  y = mpg,
                  label = c(round(car_0,2), round(car_1,2)), fill = factor(mpg)), show.legend = FALSE,
              color = "white",
              fontface = "bold"
              ) + 
    
    # Adding title
    labs(title = "Regression Lines by Vehicle Transmission",
         subtitle = "mpg outcome based on the average for each type of vehicle transmission") + 
    
    # Editing the x-axis
    xlab(label = "Weight [ton]") + 
    
    # Editing the y-axis
    ylab(label = "Miles per Gallon [mpg]") + 
    
    # Moving the legend to bottom.
    theme(legend.position = "bottom",
          legend.title.align = 0.5,
          plot.title = ggplot2::element_text(hjust = 0.5)) +
    
    # Removing the legend title.
    ggplot2::guides(fill = ggplot2::guide_legend(title = ""))
```
\center

**Figure 4** -- On average, a manual car yields 7.24 more miles per gallon than automatics.